### 2012-02-13 Text/paper.tex
First import.

\documentclass{article}
\title{Binary fluid simulations with complicated geometries}
\begin{document}
\maketitle
We will start description here.
\end{document}