### 2012-02-16 Text/paper.tex
Minor changes.

\documentclass{article}
\title{Binary fluid simulations with complicated geometries}
\begin{document}
\maketitle
Need to start here.
\end{document}